### 2017-06-15 src/5_3_Proposal.tex
saving my work: adding background sections to thesis

VECM introduces the long-run relationship among a set of cointegrated variables
as an error correction term. VECM is a special case of the vector autorregresive
model (VAR) model. VAR model expresses future values as a linear combination of
variables past values.  However, VAR model cannot be used with non-stationary
variables, which is a common feature of financial assets. VECM is a linear model
but in terms of variable differences. If cointegration exists, variable
differences are stationary and they introduce an error correction term which
adjusts coefficients to bring the variables back to equilibrium. In finance,
many economic time series turn to be stationary when they are differentiated and
cointegration restrictions often improves forecasting~\cite{duy1998}. Therefore,
VECM has been widely adopted.

@@

Our proposal is an online formulation of the VECM called Online VECM (OVECM)
based on consideration of only a sliding window of the historical data.  OVECM
introduces matrix optimizations in order to reduce the number of operations and
also takes into account the fact that cointegration vector space doesn't
experience large changes with small changes in the input data. Moreover, OVECM
uses RR instead of OLS to obtain VECM parameters. Our method is later tested using four
currency rates from the foreign exchange market with a frequency of 10 seconds.
Model efectiveness is focused on out-of-sample forecast rather than in-sample
fitting.  This criteria allows the OVECM prediction capability to be expressed
rather than just explaining data history. Our method performance is compared
with its optimal offline algorithm.

%The next sections are organized as follows: section~\ref{sec:background}
%presents the VAR and VECM, the OVECM algorithm proposed is presented in
%section~\ref{sec:methodology}. Section~\ref{sec:results} gives a description of
%the data used and the tests carried on to show accuracy and time comparison of
%our proposal against the traditional VECM and
%section~\ref{sec:conclusions} includes conclusions and a proposal for future
%tudy.

>>>>

%So why do we care about cointegration? Someone else can probably give more
%econometric applications, but in quantitative finance, cointegration forms the
%basis of the pairs trading strategy: suppose we have two cointegrated stocks X
%and Y, with the particular (for concreteness) cointegrating relationship X - 2Y
%= Z, where Z is a stationary series of zero mean. For example, X could be
%McDonald's, Y could be Burger King, and the cointegration relationship would
%mean that X tends to be priced twice as high as Y, so that when X is more than
%twice the price of Y, we expect X to move down or Y to move up in the near
%future (and analogously, if X is less than twice the price of Y, we expect X to
%move up or Y to move down). This suggests the following trading strategy: if X -
%2Y > d, for some positive threshold d, then we should sell X and buy Y (since we
%expect X to decrease in price and Y to increase), and similarly, if X - 2Y < -d,
%then we should buy X and sell Y.

@@

In this paper, we propose an online formulation of the VECM called Onlive VECM
(OVECM) based on consideration of only a sliding window of the most recent data.
The algorithm introduces matrix optimizations in order to reduce the number of
operations and also takes into account the fact that cointegration vector space
doesn't experience large changes with small changes in the input data. Moreover,
RR instead of OLS is used to solve VECM. Our method is later tested using four
currency rates from the foreign exchange market with different frequencies.  Model
efectiveness is focused on out-of-sample forecast rather than in-sample fitting.
This criteria allows the OVECM prediction capability to be expressed rather than
just explaining data history. Our method performance is compared with its
optimal offline algorithm.

The next sections are organized as follows: section~\ref{sec:background}
presents the VAR and VECM, the OVECM algorithm proposed is presented in
section~\ref{sec:methodology}. Section~\ref{sec:results} gives a description of
the data used and the tests carried on to show accuracy and time comparison of
our proposal against the traditional VECM and
section~\ref{sec:conclusions} includes conclusions and a proposal for future
tudy.
